SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-11 docs/questioning.tex
material for discussion and lessons learned

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it etiological}, {\it symptomatic} and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
%% PL:
%% Explain what "by design" means.
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.
%% PL This paragraph repeats the material of the previous one, doesn't
%% it?

@@

This approach is used to detect {\sc xss} and {\sc csrf} attacks.
%% PL All of them more or less target web users, correct?
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
Policies can be expressed through JavaScript extensions,
pattern matching or syntax specific settings.
%% PL What kind of policies are these? Syntax-based, pattern matching?
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

@@

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Point {\sc SB}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Point {\sc ub}).
%% PL In which way is this a kind of policy enforcement?
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.

A critical issue regarding
the above frameworks is that they cannot
define policies that restrict the behavior of third-party
scripts introduced by script inclusion and {\tt iframe}
tags. For instance, consider a web page (foo.net)
that prints the value of a query parameter ({\tt query})
from the page's {\sc url} in the web page's content
without escaping the value. A malicious user
can take advantage of this and inject an {\tt iframe}
into the page to steal the user's cookie and
send it via an image request to a web site
that he or she controls (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

\noindent
{\it WebJail}~\cite{VDDPJ11}, and
{\sc soma}~\cite{OWVS08} are two frameworks that
can actually detect such an attack (Point {\sc ub}).
To achieve this, {\sc soma} requires site administrators
to specify legitimate, external domains for sending
or receiving information in order to approve interactions
between them and the protected web site. As a result,
{\sc soma} can also detect {\sc csrf} attacks.
WebJail contains the functionality of third-party scripts
by introducing a web component integrator that restricts
the access that these scripts may have to either the data
or the functionality of other components.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc eq}).
% However, this is a simplistic approach that may
% lead to many false positives.

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc eq}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc eq}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
% Note that the above schemes detect potential attacks
% in locus {\sc z}. This automatically means that they
% cannot detect non-persistent JavaScript-driven
% {\sc xss} attacks. Although they can be
% modified to do so.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc sb}).

@@

\subsubsection{Training}

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly}
do not tag transactions with their corresponding call-sites.
In this case a false negative result
is prominent. For instance, consider an application that will send the
password for a forgetful user via email by executing the following
query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit} and {\it libAnomaly}, could be easily escaped
since a malicious user, could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent
%% PL I do not understand what is specific to training here; most
%% classic SQL injection attacks work in this way.
However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\cite{APKLM10}.
%% PL Like the return-to-lib attack?

@@

All the above mechanisms are relatively easy to deploy.
{\sc dsl} injection attack countermeasures
can be retrofitted to a system typically by changing
some configuration files. This does not apply
to {\sc amnesia} though, since significant source code
modifications are required for every query that exists
in the application. Finally, {\sc swap} and {\sc xssds}
are implemented within a proxy on the server-side.

@@

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{defect-percentages.png}
\end{center}
\caption{\label{fig:defect_sources}A pie showing the sources
where the authors reffered to, to find vulnerabilities to test
their mechanisms. We also provide which source was used
in which publication:
{\sc cve}~\cite{XBS06,NLC07,VFJKKV07,PMP11,BK04,JB07,SMS13,WPLKK09,JKK06a},
{\it Bugtraq}~\cite{PB05,KKVJ06,KJKV09,JEP08},
{\sc xss}ed.com~\cite{NSS06,APKLM10},
ha.ckers.org~\cite{TNH07,PSC09,LV09},
Microsoft~\cite{RDWDE07}.}
\end{figure}

@@

\subsection{Ease of Use}

As we mentioned in Subsection~\ref{sec:deployment},
there are various reasons why a mechanism could
be difficult to deploy.

Consider the majority of mechanisms coming from
the policy enforcement subcategory. In most cases
developers should modify multiple components
to use each mechanism. Specifically, mechanisms like
BrowserShield, and {\sc beep} require modifications
both on the server and the client side.
Thus, it would be difficult for them to be adopted by
both browser vendors and application developers.
On the other hand, there are cases where the policies
introduced in the server, are enforced on the client-side,
via a library embedded in the server's response
(i.e. Blueprint). Such an approach is really convenient
since no modifications are needed on the client.

Extensive modifications in the application's source
code can also be a reason that can make a mechanism
difficult to use. {\sc sql}rand, {\sc amnesia}
and mechanisms coming from the taint tracking
subcategory are such examples.
In the first two cases, programmers should modify every
code fragment that involves the execution of a query.
In the latter case they should also change
all the code fragments that involve user-input handling. 
In addition, there are cases (i.e. ConScript and the system
by Phung et al.) that require many modifications
by the developers in order to introduce their policies.
Note also that in many cases, developers are required to
learn a new {\sc api}.

Finally, a disadvantage of mechanisms that involve training
({\sc sd}river, {\sc xss-guard}), is that when code statements
are altered, a new training phase is necessary. However,
with the increased adoption of the use of automated testing
and continuous integration frameworks, this phase could be
easily repeated.

@@

@@

In frameworks that deal with JavaScript-driven {\sc xss} attacks we
see that attack detection may take place either on the server-side or
on the client-side. In the first case, a proxy is placed on the server
to examine the server's responses before they reach the user's web
browser. In the second case a modified browser or a library that is
securely downloaded from the server checks the responses for potential
attacks. We find that the tendency so far is to create frameworks that
perform detection on the client side: in particular, 11 out of 19
frameworks (57,8\%) detect JavaScript-driven {\sc xss} attacks at the
client's side. Note that the majority of the mechanisms that detect
such attacks on the server-side can also detect {\sc dsl} code
injection attacks (7 out of 8).

@@

Equations~\ref{eq:ppv-se-sp} and~\ref{eq:npv-se-sp} use true and false
positives and negatives, like equations~\ref{eq:sensitivity}
and~\ref{eq:specificity}. However, \textsc{tp}, \textsc{tn},
\textsc{fp}, and \textsc{fn} are not the same in the two cases:
whereas sensitivity and specificity are measured on a test
environment, \textsc{ppv} and \textsc{npv} are measured on the real,
target environment. In fact, if \textsc{pr} is the probability of
attacks in the real world, the prevalence, then we
have~\cite{linn2004,altman1994}:

@@

The prevalence is the prior probability that an event is an attack,
based on our understanding of the volume and frequency of attacks; the
\textsc{ppv} and \textsc{npv} are the revised estimates of the
probability based on the results of the detection
mechanism~\cite{altman1994}. The lower the prevalence of an attack,
the more confident we can be that a negative test result indicates
that no attack has taken place and the less sure we can be that a
positive test result indicates a real attack.

@@

Another interesting finding involves the computational overhead.
We observe that the average overhead that can be extracted from
the publications that provide the overhead in percentage terms is
18\%. Relative research on protection mechanisms~\cite{SPWS13}
has indicated that mechanisms that introduce an overhead
larger than 10\% do not tend to gain wide adoption in
production environments which in this case could be a
reason why some these tools have not been adopted.

Cite~\cite{DKH14}.

We believe that our research may facilitate researchers
or practitioners who would like to use one of the
aforementioned mechanisms and researchers who would like
to build new ones. Especially in the second case,
researchers can utilize our exploitation model as a
starting point to develop their approach.

@@

We believe that forming a
standard way to test such mechanisms in terms
of accuracy and computational performance
would be one of the reasons of adopting them.

>>>>

Figure~\ref{fig:defenses} presents a taxonomy of the
defenses that protect web applications against attacks based on
code injection vulnerabilities.
We can identify three approaches:
{\it etiological}, {\it symptomatic} and {\it hybrid}.
The etiological category involves mechanisms designed to
stop attacks based on their causes and origins~\cite{JL75,L81}. 
The symptomatic category incorporates a variety of schemes that
inspect the behavior of applications and detect attacks based on
their undesirable symptoms~\cite{D76,A00}.
Hybrid mechanisms borrow characteristics from both
categories. Table~\ref{tab:comp} groups the various subcategories and
for every mechanism provides the following:

@@

Parse-tree validation seems to be an efficient approach
to detect {\sc dsl} code injection attack,
especially in the case of the {\sc sql}check
implementation. This however does not apply in
the case of the mechanisms that borrow elements from
this approach and examine the syntax trees
of scripts to detect JavaScript-driven {\sc xss} attacks
as we will see in Subsection~\ref{sec:hybrid}.

@@

This approach is used to detect {\sc xss} and {\sc csrf} attacks.
When using a framework that implements
this approach, developers must define
specific security policies on the server side.
Policies can be expressed through JavaScript extensions,
pattern matching or syntax specific settings.
Then the policies are enforced either in the user's
browser at runtime or in a server-side proxy that intercepts
server responses.

@@

Another policy enforcement approach introduces
policies directly either in {\sc html} or JavaScript code
to confine their behavior. {\it BrowserShield}~\cite{RDWDE07}
acts as a proxy on the server side (Point {\sc SB}) to
parse the {\sc html} of server responses and identifies
scripts. Then, it rewrites them into safe equivalents
and protects the web user from exploits
that are based on reported browser vulnerabilities.
An issue regarding this framework (as well as {\sc beep}
and {\sc dsi}) is that it cannot define policies that
restrict the behavior of third-party scripts introduced
by script inclusion and {\tt iframe} tags.
For instance, consider a web page (foo.net)
that prints the value of a query parameter ({\tt query})
from the page's {\sc url} in the web page's content
without escaping the value. A malicious user
can take advantage of this and inject an {\tt iframe}
into the page to steal the user's cookie and
send it via an image request to a web site
that he or she controls (malicious.com).
This could be achieved by including the following
link to the malicious web site (or send it via phishing
email) and induce the user to click on it:

@@

\noindent
The following frameworks can actually detect such
an attack (Point {\sc ub}).
{\it ConScript}~\cite{ML10}, {\it CoreScript}~\cite{YCIS07}
and the framework by {\it Phung et al.}~\cite{PSC09}
extend JavaScript with new primitive functions that
provide safe methods to protect potentially vulnerable
JavaScript functions. In both cases, policy enforcement takes
place on the JavaScript engine of the browser (Point {\sc ub}).
In this way, {\sc xxs} attacks that take advantage
functions like {\tt write} and {\tt eval}, which are
used to assemble innocuous-looking parts into harmful
strings,\footnote{The infamous Sammy worm that
infected MySpace in 2005~\cite{SP07,ELX07}
utilized the {\tt eval} function to assemble a
script with malicious behavior.} would fail.
{\it WebJail}~\cite{VDDPJ11}, is a framework similar to
ConScript. WebJail, contains the functionality of
third-party scripts by introducing a web component
integrator that restricts the access that these scripts
may have to either the data or the functionality of other components
Finally, {\sc soma}~\cite{OWVS08} requires site administrators
to specify legitimate, external domains for sending
or receiving information in order to approve interactions
between them and the protected web site. As a result,
{\sc soma} can also detect {\sc csrf} attacks.

@@

There are different implementations of this approach
in terms of how the tainted data is marked and observed,
and how attacks are detected.
For example, {\it Haldar et al.}~\cite{HCF05} have implemented
their scheme for the Java Virtual Machine ({\sc jvm})
where they instrument various classes. When a
tainted string is used as an argument to a sink method
then an exception is raised (Point {\sc eq}).

It is possible to apply further checks when it is established that
tainted data have reached a sink (Point {\sc eq}). {\it Xu et
  al.}~\cite{XBS06} track taint information at the level of bytes in
memory. To distinguish between legitimate and malicious uses of
untrusted data that reach a sink, they search the data for suspicious
symbols by using regular expressions. {\sc csse}~\cite{PB05}
associates tainted data with specific metadata. Such metadata include
the origins of tainted data, its propagation within the application,
and others. When tainted data reaches a sink, {\sc csse} performs
syntactic checks based on the aforementioned metadata (Point {\sc eq}).
{\it {\sc php} Aspis}~\cite{PMP11} works in a similar way. To obtain
metadata, it takes advantage of {\sc php} arrays data structure.
Finally, {\sc wasc}~\cite{NLC07} analyzes {\sc html} responses to check
if there are tainted data that contain scripts (Point {\sc sb}).

@@

\subsubsection{Training}
\label{sec:train}

@@

Defenses based on training include some mechanisms that can be easily
circumvented. For example, {\sc didafit} and {\it libAnomaly}
do not tag transactions with their corresponding call-sites.
In this case a false negative result
is prominent. For instance, consider an application that will send the
password for a forgetful user via email by executing the following
query:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice'
\end{lstlisting}
\end{small}
\egroup
\noindent
This same application could allow users to lock their terminal,
but allow the unlocking either with the user's password or with
the administrator password (the 4.3 {\sc bsd} {\em lock} command
behaved in this peculiar way).
The corresponding query to verify the password on the locked
workstation would be as follows:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
SELECT password from userdata WHERE id = 'Alice' OR
id = 'admin'
\end{lstlisting}
\end{small}
\egroup
\noindent
In this case {\sc didafit} and {\it libAnomaly}, could be easily escaped
since a malicious user, could obtain the administrator's password
by email by entering on the password retrieval form
as his user identifier the following:
\bgroup
\lstset{language=SQL}
\begin{small}
\begin{lstlisting}
nosuchuser' OR id = 'admin
\end{lstlisting}
\end{small}
\egroup
\noindent

However, mechanisms that create signatures that involve
elements not only associated with the code statements
(i.e. {\sc amnesia} and {\sc sd}river)
could detect such attacks. Specifically,
{\sc sd}river associates a complete stack trace with the
root of an {\sc sql} statement, thus it can correlate
queries with their call sites and detect attacks
like the aforementioned one.
Furthermore, training tools that detect
{\sc xss} attacks based on JavaScript injection
would fail to detect a return-to-JavaScript
attack~\cite{APKLM10}.

@@

Most of the mechanisms above are relatively easy to deploy.
{\sc dsl} injection attack countermeasures
can be retrofitted to a system typically by changing
some configuration files (i.e. {\sc sd}river). This does not apply
to {\sc amnesia} though, since significant source code
modifications are required for every query that exists
in the application. Finally, {\sc swap} and {\sc xssds}
are implemented within a proxy on the server-side.

@@

\begin{figure}
\begin{center}
\leavevmode
\includegraphics[scale=0.47]{defect-percentages.png}
\end{center}
\caption{\label{fig:defect_sources}A pie showing the sources
where the authors reffered to, to find vulnerabilities to test
their mechanisms. We also provide which source was used
in which publication:
{\sc cve}~\cite{XBS06,NLC07,VFJKKV07,PMP11,BK04,BV08,JB07,SMS13,WPLKK09,JKK06a,PS11},
{\it Bugtraq}~\cite{PB05,KKVJ06,JEP08},
{\sc xss}ed.com~\cite{NSS06,APKLM10},
ha.ckers.org~\cite{TNH07,PSC09,LV09},
Microsoft~\cite{RDWDE07}.}
\end{figure}

@@

\subsection{Ease of Use}
\label{sec:deploy2}

As we observed in Subsection~\ref{sec:deployment},
there are various reasons why a mechanism could
be difficult to deploy.

Consider the majority of mechanisms coming from
the policy enforcement subcategory. In most cases
developers should modify multiple components
to use each mechanism. Specifically, mechanisms like
BrowserShield, and {\sc beep} require modifications
both on the server and the client side.
Thus, it would be difficult for them to be adopted by
both browser vendors and application developers.
On the other hand, there are cases where the policies
introduced in the server are enforced on the client-side,
via a library embedded in the server's response
(i.e. Blueprint). Such an approach is really convenient
since no modifications are needed on the client.

Extensive modifications in the application's source
code can also be a reason that can make a mechanism
difficult to use. {\sc sql}rand, {\sc amnesia},
mechanisms coming from the parse-tree validation
subcategory and mechanisms coming from the taint tracking
subcategory are such examples.
In the first three cases, programmers should modify every
code fragment that involves the execution of a query.
In the latter case they should also change
all the code fragments that involve user-input handling. 
In addition, there are cases (i.e. ConScript and the system
by Phung et al.) that require many modifications
by the developers in order to introduce their policies.

Finally, even though many of the mechanism that involve
training are easy to deploy, they have a distinct disadvantage.
When the application is altered, mechanisms like {\sc sd}river
and {\sc xss-guard} require a new training phase is necessary.
However, with the increased adoption of automated testing
and continuous integration frameworks, this phase could be
easily repeated.

@@

While describing the various mechanisms in
Section~\ref{sec:defs}, we observed that some
can be bypassed by attackers that know how they
work.

However, some mechanisms are designed in such a way
so that they can be extended and become immune to
the attacks that they are currently vulnerable to. For
example, policy enforcement frameworks that are
based on JavaScript or {\sc html} rewriting
can be extended to detect attacks that leverage
{\tt iframe} tags (see Subsubsection~\ref{sec:policy}).

This does not apply to all mechanisms though.
In particular, training mechanisms like {\sc didafit},
and {\sc xss-guard} must be redesigned
to detect the attacks that were described in
Subsubsection~\ref{sec:train} and
Subsection~\ref{sec:hybrid} respectively.

@@

In frameworks that deal with JavaScript-driven {\sc xss} attacks we
see that attack detection may take place either on the server-side or
on the client-side. In the first case, a proxy is placed on the server
to examine the server's responses before they reach the user's web
browser. In the second case a modified browser or a library that is
securely downloaded from the server checks the responses for potential
attacks. We find that the tendency so far is to create frameworks that
perform detection on the client side: in particular, 11 out of 19
frameworks (57,8\%) detect JavaScript-driven {\sc xss} attacks on the
client's side. Note that the majority of the mechanisms that detect
such attacks on the server-side can also detect {\sc dsl} code
injection attacks (7 out of 8).

@@

\noindent
Equations~\ref{eq:ppv-se-sp} and~\ref{eq:npv-se-sp} use true and false
positives and negatives, like equations~\ref{eq:sensitivity}
and~\ref{eq:specificity}. However, \textsc{tp}, \textsc{tn},
\textsc{fp}, and \textsc{fn} are not the same in the two cases:
whereas sensitivity and specificity are measured on a test
environment, \textsc{ppv} and \textsc{npv} are measured on the real,
target environment. In fact, if \textsc{pr} is the probability of
attacks in the real world, the prevalence, then we
have~\cite{linn2004,altman1994}:

@@

\noindent
The prevalence is the prior probability that an event is an attack,
based on our understanding of the volume and frequency of attacks; the
\textsc{ppv} and \textsc{npv} are the revised estimates of the
probability based on the results of the detection
mechanism~\cite{altman1994}. The lower the prevalence of an attack,
the more confident we can be that a negative test result indicates
that no attack has taken place and the less sure we can be that a
positive test result indicates a real attack.

@@

Moreover, extensive accuracy tests may lead to more
well-designed and robust mechanisms. Observe that mechanisms
that can be circumvented were not extensively tested
in terms of accuracy. For instance, {\sc didafit}
was not tested at all and {\sc xss-guard}'s testing involved
6 known attacks. This also applies to some frameworks coming
from the taint tracking subcategory.
Perhaps with more tests the authors would design
their mechanisms differently.

Limiting the scope of the attacks to
be detected may lead to an efficient tool.
Consider the system by Stock et al. It is the only one
that targets exclusively {\sc dom}-based
{\sc xss} attacks and detects them in an accurate manner
as seen in Table~\ref{tab:comp}. Notice also
its extensive testing in real-world attacks.

Another interesting finding involves the computational overhead.
We observe that the average overhead that can be extracted from
the publications that provide the overhead in percentage terms is
18\%. Relative research on protection mechanisms~\cite{SPWS13}
indicates that mechanisms that introduce an overhead
larger than 10\% do not tend to gain wide adoption in
production environments. Hence, the computational overhead
of some mechanisms could be a reason why they have not been adopted.

We need to mention that the fact that mechanisms come
sometimes with deployment difficulties does
not necessarily mean that this is a serious reason
not to adopt them. The deployment of such frameworks
is most of the time complex because as we observed
in Section~\ref{sec:attacks} the attacks that they
detect are complex too. So every security feature can
come with such a trade-off and those who intend to use it
should be aware of this.

Deployment can also depend on the detection point
of an attack. We saw that many policy enforcement
mechanisms that detect attacks at Point {\sc ub},
are not easy to deploy because modifications
are needed both on the server and the client side.
We observed though that Blueprint has not such
deployment issues (see Subsection~\ref{sec:deploy2}).
Thus, when developing a new countermeasure,
researchers should consider where they are planning
to detect attacks, observe the corresponding
deployment challenges, and try to mitigate them.

% discuss Availability. Finally,

@@

\begin{itemize}
\item Cite~\cite{DKH14}.
\item We believe that our research may facilitate researchers
or practitioners who would like to use one of the
aforementioned mechanisms and researchers who would like
to build new ones. Especially in the second case,
researchers can utilize our exploitation model as a
starting point to develop their approach.
\item We believe that forming a
standard way to test such mechanisms in terms
of accuracy and computational performance
would be one of the reasons of adopting them.
\end{itemize}
